---
title: "Modeling the role of negative affect in the relationship between everyday pain and working memory"
author: Steven R. Anderson(a), Joanna E. Witkin(a), Taylor Bolt, Maria M. Llabre, Elizabeth A. Reynolds Losin; (a) Indicates co-first authorship

output:
  html_document:
    fig_width: 6
    fig_height: 4
    toc: true
    toc_float: true
    smooth_scroll: true
    theme: united
    highlight: tango
---
<style type="text/css">

a:link {
    color: DarkBlue;
}

body{ /* Normal  */
   font-size: 14px;
}
td {  /* Table  */
   font-size: 12px;
}
h1 { /* Header 1 */
 font-size: 26x;
 color: DarkBlue;
}
h2 { /* Header 2 */
 font-size: 26px;
 color: DarkBlue;
}
h3 { /* Header 3 */
 font-size: 18px;
 color: DarkBlue;
}
h4 { /* Header 4 */
 font-size: 22px;
 color: DarkOrange;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

#Setup
```{r setup, message=FALSE, warning=FALSE, echo=FALSE,cache=FALSE}

library(lavaan)
library(foreign)
library(plyr)
library(dplyr)
library(reshape2)
library(lattice)
library(table1)#descriptive statistics table
library(tidyr)
library(semPlot)
library(corrplot)
library(kutils)
library(psych)
library(moments)#tests skewness and kurtosis
library(ggplot2)
library(semTable)
library(ggpubr)
library(influence.SEM)
library(ggrepel)
library(faoutlier)

#disable scientific notation
options(scipen=999)

rm(list=ls()) #removes all the variables from workspace

require("knitr")
opts_knit$set(root.dir = "//datastore01.psy.miami.edu/Groups/ELosin_Lab/Research/Research_Projects/SEM_PainCognition/Data_Files")
```

#Data Preparation
```{r Data Preparation, echo=FALSE, message=FALSE, cache=FALSE}
#RAW HCP BEHAVIORAL DATA
HCP_behav_data1 <- read.csv("HCP1200_Restricted_Behav.csv")
HCP_behav_data2 <- read.csv("HCP1200_Unrestricted_Behav.csv")

#subset only age in years and subject ID from restricted data
myvars <- c("Subject", "Age_in_Yrs","Race","Ethnicity")

HCP_behav_data1.5 <- HCP_behav_data1[myvars]

#merge subsetted restricted and unrestricted data
HCP_behav_data_all <- merge(HCP_behav_data1.5,HCP_behav_data2,by="Subject")

#LIST OF SUBJECTS WE HAVE FMRI DATA FOR
subject_list <- read.csv('SubjectList_HCP416.csv',header=F) #these are the subjects we have fMRI data for
colnames(subject_list)[1] <- "Subject"

#fMRI WM task related activation 
WM_ROI <- read.csv("WM_all_ROI.csv") 

#RAW HCP PAIN INTENSITY DATA (EXCLUDED FROM MAIN DF DUE TO TOOLBOX GLITCH; RECEIVED VIA EMAIL FROM HCP)
HCP_painintensity <- read.csv("S1200_TB_Pain_Intensity_Raw.csv") 
colnames(HCP_painintensity)[3] <- "Pain_Intensity"

#combine subject list of subjects we have fmri data for with full HCP 1200 behavioral dataframe
HCP2 <- merge(subject_list,HCP_behav_data_all,by="Subject")

#combine pain intensity variable with fmri subject list
HCP3 <- merge(HCP_painintensity,subject_list,by="Subject",all.y=T)

#identify suplicated Subject IDs due to test-retest of pain intensity
#duplicated(HCP3$Subject)
painint_duplicates<-as.data.frame(HCP3$Subject[duplicated(HCP3$Subject)]) #20 subjects are affected

#NO DUPLICATES HCP PAIN INTENSITY DATA (EXCLUDED FROM MAIN DF DUE TO TOOLBOX GLITCH; RECEIVED VIA EMAIL FROM HCP)
HCP_painintensity2 <- read.csv("S1200_TB_Pain_Intensity_Raw_firstscores.csv")
colnames(HCP_painintensity2)[3] <- "Pain_Intensity"

#combine pain intensity variable (no duplicates) with fmri subject list
HCP3.2 <- merge(HCP_painintensity2,subject_list,by="Subject",all.y=T)

#combine raw 416 behavioral data with pain intensity variable
HCP4 <- merge(HCP2,HCP3.2,by="Subject")

#combine merged behavioral data with WM ROI data
SEM <- merge(HCP4,WM_ROI,by="Subject")

#examine data
#names(SEM)
#head(SEM)

#examine pain intensity variable
summary(SEM$Pain_Intensity)
str(SEM$Pain_Intensity)

#rename variables of interest to easier to type names
SEM$WMAcc_2bk <- SEM$WM_Task_2bk_Acc
SEM$WMAcc_0bk <- SEM$WM_Task_0bk_Acc
SEM$Sadness <- SEM$Sadness_Unadj
SEM$Lonely <- SEM$Loneliness_Unadj
SEM$PercHost <- SEM$PercHostil_Unadj
SEM$PercRej <- SEM$PercReject_Unadj
SEM$Fear <- SEM$FearAffect_Unadj
SEM$Anger <- SEM$AngAffect_Unadj
SEM$Hostil <- SEM$AngHostil_Unadj
SEM$Aggr <- SEM$AngAggr_Unadj
SEM$FearBody <- SEM$FearSomat_Unadj
SEM$PainInterf<- SEM$PainInterf_Tscore 
SEM$Stress <- SEM$PercStress_Unadj

#----------------------------------------------------
#define ROIs

#get mean of the following parcels to create a larger dMFC that more closely covers the amount of brain area 

SEM$dMFC <- rowMeans(SEM[c("ROI_4","ROI_23","ROI_24","ROI_25","ROI_34")], na.rm=TRUE)
#Gordon parcel categories: ROI_4=Default, ROI_23=VentralAttn, ROI_24=FrontoParietal, ROI_25=Default, ROI_34=CinguloOperc

SEM$aMCC <- rowMeans(SEM[c('ROI_22', 'ROI_27')], na.rm=TRUE)
#Gordon parcel categories: ROI_22=CinguloOperc, ROI_27=CinguloOperc

SEM$vmPFC <- rowMeans(SEM[c('ROI_116', 'ROI_117')], na.rm=TRUE)
#Gordon parcel categories: ROI_116=Default, ROI_117=Default

#-----------------------------------------
#PARTICIPANT CHARACTERISTICS 

#calculate age stats 
summary(SEM$Age_in_Yrs) #Mean = 28.59
sd(SEM$Age_in_Yrs) #SD = 3.72

#count number of males and females
str(SEM$Gender)
table(SEM$Gender) #216 female

#count race
str(SEM$Race)
table(SEM$Race)
# Am. Indian/Alaskan Nat.  1 
# Asian/Nat. Hawaiian/Othr Pacific Is. 30 
# Black or African Am. 61 
# More than one 13
# Unknown or Not Reported 8
# White 303 

#count ethnicity
str(SEM$Ethnicity)
table(SEM$Ethnicity)
# Hispanic/Latino   43  
# Not Hispanic/Latino 370
# Unknown or Not Reported 3

#subset Non-Hispanic White participants
nhw <- SEM[ which(SEM$Race=='White' & SEM$Ethnicity=='Not Hispanic/Latino'),] #271/416 = 

#-----------------------------------------
#subset only variables you need
SEM2<-SEM[c(1:7,112,162,442:495,570,578,912:927)]

#make pain intensity numeric 
SEM2$Pain_Intensity<-as.numeric(SEM2$Pain_Intensity)

```

#Create zero pain subset
```{r Create zero pain subset, echo=FALSE, message=FALSE, cache=FALSE}
#create subset of subjects with non-zero pain intensity 
SEM2_nonzero_painint <- SEM2[ which(SEM2$Pain_Intensity > 0), ]

```

#Examine non-zero pain intensity subset
```{r Examine non-zero pain intensity subset, echo=FALSE, message=FALSE, cache=FALSE}
 #examine data
#histogram full sample
histogram(~ Pain_Intensity, data = SEM2, type = "count",col="royalblue1",xlab="Pain Intensity",main="Full Sample")

#histogram nonzero pain intensity subset
histogram(~ Pain_Intensity, data = SEM2_nonzero_painint, type = "count",col="royalblue1",xlab="Pain Intensity",main="Non-Zero Pain Subset")

#density plot - full sample
ggdensity(SEM2$Pain_Intensity, color="gray18",fill="royalblue1",
          xlab = "Pain Intensity",alpha=1,main="Full Sample")

#density plot - non-zero pain subset
ggdensity(SEM2_nonzero_painint$Pain_Intensity,fill="royalblue1",main = "Non-Zero Pain Subset",alpha=1,
          xlab = "Pain Intensity")

```

#Outliers
```{r Outliers, echo=FALSE, message=FALSE, cache=FALSE}
#cooks distance method for detecting multivariate outliers
paintintsmod_lm <- lm(WMAcc_2bk ~ Pain_Intensity + vmPFC + dMFC + aMCC + Sadness + Fear + Anger + Stress, data=SEM2)
cooksd <- cooks.distance(paintintsmod_lm)

#plot cook's d
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 1, col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>1,names(cooksd),""), col="red")  # add labels

#extract individual rows
influential <- as.numeric(names(cooksd)[(cooksd > 1)])  # influential row numbers

#cooks distance method for detecting multivariate outliers (cutoff 1)
#plot cook's d
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 1, col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>1,names(cooksd),""), col="red")  # add labels

```

#Sample Characteristics Table
```{r Sample Characteristics Table, echo=FALSE, message=FALSE, cache=FALSE}
SEM2b<-SEM2_nonzero_painint

#Sample Characteristics Table with table1 package
table1::label(SEM2b$Age_in_Yrs) <- "Age"
table1::label(SEM2b$Race) <- "Race"
table1::label(SEM2b$Ethnicity) <- "Ethnicity"
SEM2b$Gender <- 
  factor(SEM2b$Gender, levels=c("F","M"),
         labels=c("Female", 
                  "Male"))
table1::table1(~Age_in_Yrs + Race + Ethnicity + Gender, data = SEM2b)

```

#Descriptive Statistics Table
```{r Descriptive Statistics Table, echo=FALSE, message=FALSE, cache=FALSE}
#subset only model variables
SEM2_nonzero_painint_sub<-SEM2_nonzero_painint[c(8,65,9,66,68,72,73,77,78,79:81)]

#Descriptive Statistics Table with table1 package
table1::label(SEM2_nonzero_painint_sub$Pain_Intensity) <- "Pain Intensity"
table1::label(SEM2_nonzero_painint_sub$WMAcc_2bk) <- "2-Back Accuracy"
table1::label(SEM2_nonzero_painint_sub$ListSort_AgeAdj) <- "List Sort Score (Age Adjusted)"
table1::label(SEM2_nonzero_painint_sub$Anger) <- "NIH Toolbox Anger-Affect Survey"
table1::label(SEM2_nonzero_painint_sub$Stress) <- "NIH Toolbox Perceived Stress Survey"
table1::label(SEM2_nonzero_painint_sub$Sadness) <- "NIH Toolbox Sadness Survey"
table1::label(SEM2_nonzero_painint_sub$Fear) <- "NIH Toolbox Fear-Affect"
table1::label(SEM2_nonzero_painint_sub$aMCC) <- "aMCC Activity"
table1::label(SEM2_nonzero_painint_sub$dMFC) <- "dMFC Activity"
table1::label(SEM2_nonzero_painint_sub$vmPFC) <- "vmPFC Activity"

#table with only variables included in SEM
table1::table1(~Pain_Intensity + WMAcc_2bk + ListSort_AgeAdj + Anger + Stress + Sadness + Fear + aMCC + dMFC + vmPFC, data = SEM2_nonzero_painint_sub) 

```

#Diagnostics
```{r Data Diagnostics, echo=FALSE, message=FALSE, warning=FALSE,cache=FALSE}
#assess normality, skewness, and kurtosis of each variable in models

#WMAcc_2bk
#descriptives
summary(SEM2_nonzero_painint_sub$WMAcc_2bk)
sd(SEM2_nonzero_painint_sub$WMAcc_2bk,na.rm=T)

#density plot
ggdensity(SEM2_nonzero_painint_sub$WMAcc_2bk, 
          main = "Density plot of WMAcc",
          xlab = "WMAcc_2bk")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint_sub$WMAcc_2bk)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint_sub$WMAcc_2bk)

#histogram
hist(SEM2_nonzero_painint_sub$WMAcc_2bk)

#skewness (values >= 2 indicates problem)
skewness(SEM2_nonzero_painint_sub$WMAcc_2bk,na.rm=T)

#kurtosis (values >= 7 indicates problem)
kurtosis(SEM2_nonzero_painint$WMAcc_2bk,na.rm=T)

#----------------------------------------
#ListSort_AgeAdj
#descriptives
summary(SEM2_nonzero_painint_sub$ListSort_AgeAdj)
sd(SEM2_nonzero_painint_sub$ListSort_AgeAdj,na.rm=T)

#density plot
ggdensity(SEM2_nonzero_painint_sub$ListSort_AgeAdj, 
          main = "Density plot of ListSort_AgeAdj",
          xlab = "ListSort_AgeAdj")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint_sub$ListSort_AgeAdj)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint_sub$ListSort_AgeAdj)

#histogram
hist(SEM2_nonzero_painint_sub$ListSort_AgeAdj)

#skewness (values >= 2 indicates problem)
skewness(SEM2_nonzero_painint_sub$ListSort_AgeAdj,na.rm=T)

#kurtosis (values >= 7 indicates problem)
kurtosis(SEM2_nonzero_painint$ListSort_AgeAdj,na.rm=T)

#----------------------------------------
#Pain Intensity
#descriptives
summary(SEM2_nonzero_painint$Pain_Intensity)
sd(SEM2_nonzero_painint$Pain_Intensity,na.rm=T)

#density plot - full sample
ggdensity(SEM2_nonzero_painint$Pain_Intensity, color="gray18",fill="#E7B800",alpha = .4,
          main = "Density Plot of Pain Intensity",
          xlab = "Pain Intensity")

#density plot - non-zero pain subset
ggdensity(SEM2_nonzero_painint$Pain_Intensity, color="gray18",fill="#E7B800",alpha = .4,
          main = "Density Plot of Pain Intensity",
          xlab = "Pain Intensity")

#kernel density plot 
plot(density(SEM2_nonzero_painint_sub$Pain_Intensity,na.rm=T))

#Q-Q plot
ggqqplot(SEM2_nonzero_painint_sub$Pain_Intensity)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint_sub$Pain_Intensity)

#histogram
hist(SEM2_nonzero_painint_sub$Pain_Intensity)

#skewness (values >= 2 indicates problem)
skewness(SEM2_nonzero_painint_sub$Pain_Intensity,na.rm=T)

#kurtosis (values >= 7 indicates problem)
kurtosis(SEM2_nonzero_painint_sub$Pain_Intensity,na.rm=T)

#----------------------------------------
#Pain Interference
#descriptives
summary(SEM2_nonzero_painint$PainInterf)
sd(SEM2$PainInterf,na.rm=T)

#density plot
ggdensity(SEM2$PainInterf, color="gray18",fill="#00AFBB",alpha = .4,
          main = "Density Plot of Pain Interference",
          xlab = "Pain Interference",na.rm=T)

#kernel density plot 
plot(density(SEM2$PainInterf,na.rm=T))

#Q-Q plot
ggqqplot(SEM2$PainInterf)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2$PainInterf)

#histogram
hist(SEM2$PainInterf)

#skewness 
skewness(SEM2$PainInterf,na.rm=T)

#kurtosis
kurtosis(SEM2$PainInterf,na.rm=T)

#----------------------------------------
#vmPFC
#descriptives
summary(SEM2_nonzero_painint$vmPFC)
sd(SEM2_nonzero_painint$vmPFC)

#density plot
ggdensity(SEM2_nonzero_painint$vmPFC, 
          main = "Density plot of vmPFC",
          xlab = "vmPFC")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint$vmPFC)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint$vmPFC)

#histogram
hist(SEM2_nonzero_painint$vmPFC)

#skewness 
skewness(SEM2_nonzero_painint$vmPFC,na.rm=T)

#kurtosis
kurtosis(SEM2_nonzero_painint$vmPFC,na.rm=T)
#----------------------------------------
#aMCC
#descriptives
summary(SEM2_nonzero_painint$aMCC)
sd(SEM2_nonzero_painint$aMCC)

#density plot
ggdensity(SEM2_nonzero_painint$aMCC, 
          main = "Density plot of aMCC",
          xlab = "aMCC")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint$aMCC)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint$aMCC)

#histogram
hist(SEM2_nonzero_painint$aMCC)

#skewness 
skewness(SEM2_nonzero_painint$aMCC,na.rm=T)

#kurtosis
kurtosis(SEM2_nonzero_painint$aMCC,na.rm=T)
#----------------------------------------
#dMFC
#descriptives
summary(SEM2_nonzero_painint$dMFC)
sd(SEM2_nonzero_painint$dMFC)

#density plot
ggdensity(SEM2_nonzero_painint$dMFC, 
          main = "Density plot of dMFC",
          xlab = "dMFC")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint$dMFC)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint$dMFC)

#histogram
hist(SEM2_nonzero_painint$dMFC)

#skewness 
skewness(SEM2_nonzero_painint$dMFC,na.rm=T)

#kurtosis
kurtosis(SEM2_nonzero_painint$dMFC,na.rm=T)
#----------------------------------------
#Anger
#descriptives
summary(SEM2_nonzero_painint$Anger)
sd(SEM2_nonzero_painint$Anger,na.rm=T)

#density plot
ggdensity(SEM2_nonzero_painint$Anger, 
          main = "Density plot of Anger",
          xlab = "Anger")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint$Anger)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint$Anger)

#histogram
hist(SEM2_nonzero_painint$Anger)

#skewness 
skewness(SEM2_nonzero_painint$Anger,na.rm=T)

#kurtosis
kurtosis(SEM2_nonzero_painint$Anger,na.rm=T)

#----------------------------------------
#Stress
#descriptives
summary(SEM2_nonzero_painint$Stress)
sd(SEM2_nonzero_painint$Stress,na.rm=T)

#density plot
ggdensity(SEM2_nonzero_painint$Stress, 
          main = "Density plot of Stress",
          xlab = "Stress")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint$Stress)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint$Stress)

#histogram
hist(SEM2_nonzero_painint$Stress)

#skewness 
skewness(SEM2_nonzero_painint$Stress,na.rm=T)

#kurtosis
kurtosis(SEM2_nonzero_painint$Stress,na.rm=T)

#----------------------------------------
#Sadness
#descriptives
summary(SEM2_nonzero_painint$Sadness)
sd(SEM2_nonzero_painint$Sadness,na.rm=T)

#density plot
ggdensity(SEM2_nonzero_painint$Sadness, 
          main = "Density plot of Sadness",
          xlab = "Sadness")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint$Sadness)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint$Sadness)


#histogram
hist(SEM2_nonzero_painint$Sadness)

#skewness 
skewness(SEM2_nonzero_painint$Sadness,na.rm=T)

#kurtosis
kurtosis(SEM2_nonzero_painint$Sadness,na.rm=T)

#----------------------------------------
#Fear
#descriptives
summary(SEM2_nonzero_painint$Fear)
sd(SEM2_nonzero_painint$Fear,na.rm=T)

#density plot
ggdensity(SEM2_nonzero_painint$Fear, 
          main = "Density plot of Fear",
          xlab = "Fear")

#Q-Q plot
ggqqplot(SEM2_nonzero_painint$Fear)

#formal test of normality, p-value > .05 indicates normality
shapiro.test(SEM2_nonzero_painint$Fear)


#histogram
hist(SEM2_nonzero_painint$Fear)

#skewness 
skewness(SEM2_nonzero_painint$Fear,na.rm=T)

#kurtosis
kurtosis(SEM2_nonzero_painint$Fear,na.rm=T)

#------------------------------------------
#calculate variance inflation factor
library(car)
vif_mod<-lm(WMAcc_2bk ~ Pain_Intensity + Anger + vmPFC + aMCC + dMFC,data=SEM2_nonzero_painint)
summary(vif_mod)
vif(vif_mod)

```

#Self-reported negative affect latent construct 
```{r Self-reported negative affect latent construct , echo=FALSE, message=FALSE, cache=FALSE}
#specify model
model <- '
neg =~ Anger + Stress + Sadness + Fear 
'

#fit model
fit <- lavaan::sem(model, data=SEM2_nonzero_painint, estimator = "MLR")

#get fit indices and path estimates
summary(fit, fit.measures = TRUE, standardized=TRUE, modindices = TRUE)

#plot
semPaths(fit,what = "path",whatLabels="est",title=F,sizeMan=11,sizeLat=11,rotation=3,curvePivot = T,layout="tree",edge.label.cex=1,thresholds=TRUE,intercepts = FALSE,residuals=F,reorder=T)

```

#Zero-Order Correlations Among Variables of Interest
```{r  Zero-Order Correlations Among Variables of Interest, warnings=FALSE, echo=FALSE, message=FALSE, cache=FALSE}

#rename variables
SEM2_nonzero_painint_sub_rename<-SEM2_nonzero_painint_sub
names(SEM2_nonzero_painint_sub_rename) <- c("PSQI-Pain", "Pain Intensity","List Sorting Score","2-Back Accuracy","Sadness","Fear-Affect","Anger-Affect","Pain Interference","Perceived Stress","dMFC Activity","aMCC Activity", "vmPFC Activity")

#assess correlations in full sample
r<-corr.test(SEM2_nonzero_painint_sub_rename, y = NULL, use = "pairwise",method="pearson",adjust="fdr", alpha=.05,ci=T)

#plot matrix
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(r$r, p.mat = r$p,method = "color",col = col(200), type="full",tl.cex=.8,
         tl.col = "black", order="hclust",hclust.method="ward.D2",addrect=4, tl.srt = 45,insig = "blank",pch.col = "black")

r$r

#fdr corrected p-values 
#r$p
#-----------------------------
#assess correlations between variables of interest

#pain intensity and pain interference in full sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$PainInterf, method=c("pearson"))


#plot
ggplot(SEM2_nonzero_painint, aes(x=PainInterf, y=Pain_Intensity,color="royalblue1")) +
  geom_jitter(size=3,width=.3,color="royalblue1") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + ylab("Pain Intensity") + xlab("Pain Interference") + theme(legend.position = "none")

#pain intensity and psqi sleep in non-zero pain sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$PSQI_Pain, method=c("pearson"))


#plot
ggplot(SEM2_nonzero_painint, aes(x=PSQI_Pain, y=Pain_Intensity,color="#9475dc")) +
  geom_jitter(size=3,width=.3,color="#9475dc") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm,color="black") + theme_classic() + ylab("Pain Intensity") + xlab("PSQI - Pain") + theme(legend.position = "none") 

#-----------------------------------
#pain intensity and anger correlation in non-zero sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$Anger, method=c("pearson"))


#plot
ggplot(SEM2_nonzero_painint, aes(x=Anger, y=Pain_Intensity,color="#bf78b8")) +
  geom_jitter(size=3,width=.3,color="#bf78b8") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + ylab("Pain Intensity") + xlab("Anger")  + theme(legend.position = "none") 

#pain intensity and perceived stress correlation in non-zero sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$Stress, method=c("pearson"))


#plot
ggplot(SEM2_nonzero_painint, aes(x=Stress, y=Pain_Intensity,color="#bf78b8")) +
  geom_jitter(size=3,width=.3,color="#bf78b8") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + ylab("Pain Intensity") + xlab("Stress")  + theme(legend.position = "none") 

#pain intensity and Fear correlation in non-zero sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$Fear, method=c("pearson"))

#plot
ggplot(SEM2_nonzero_painint, aes(x=Fear, y=Pain_Intensity,color="#bf78b8")) +
  geom_jitter(size=3,width=.3,color="#bf78b8") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + ylab("Pain Intensity") + xlab("Fear")  + theme(legend.position = "none") 

#pain intensity and Sadness correlation in non-zero sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$Sadness, method=c("pearson"))

#plot
ggplot(SEM2_nonzero_painint, aes(x=Sadness, y=Pain_Intensity,color="#bf78b8")) +
  geom_jitter(size=3,width=.3,color="#bf78b8") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + ylab("Pain Intensity") + xlab("Sadness")  + theme(legend.position = "none") 

#-----------------------------------
#pain intensity and WMAcc_2bk correlation in non-zero sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$WMAcc_2bk, method=c("pearson"))

#plot
ggplot(SEM2_nonzero_painint, aes(x=WMAcc_2bk, y=Pain_Intensity,color="#de7a95")) +
  geom_jitter(size=3,width=.3,color="#de7a95") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + ylab("Pain Intensity") + xlab("2-Back Accuracy") + theme(legend.position = "none") 

#-----------------------------------
#pain intensity and ListSort correlation in non-zero sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$ListSort_AgeAdj, method=c("pearson"))

#plot
ggplot(SEM2_nonzero_painint, aes(x=ListSort_AgeAdj, y=Pain_Intensity,color="#bf78b8")) +
  geom_jitter(size=3,width=.3,color="#bf78b8") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + ylab("Pain Intensity") + xlab("List Sorting Score")  + theme(legend.position = "none") 

#---------------------------------
#ROIs and WMAcc_2bk correlation in non-zero pain
cor.test(SEM2_nonzero_painint$aMCC, SEM2_nonzero_painint$WMAcc_2bk, method=c("pearson"))

#plot 
aMCC_wmacc_plot<-ggplot(SEM2_nonzero_painint, aes(x=aMCC, y=WMAcc_2bk,color="coral")) +
  geom_point(size=2,color="coral") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + xlab("aMCC Activity") + ylab("2-Back Accuracy") + theme(legend.position = "none")

#vmPFC and WMAcc_2bk correlation in non-zero pain
cor.test(SEM2_nonzero_painint$dMFC, SEM2_nonzero_painint$WMAcc_2bk, method=c("pearson"))

#plot 
dMFC_wmacc_plot<-ggplot(SEM2_nonzero_painint, aes(x=dMFC, y=WMAcc_2bk,color="darkorchid2")) +
  geom_point(size=2,color="darkorchid2") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + xlab("dMFC Activity") + ylab("2-Back Accuracy") + theme(legend.position = "none")

#vmPFC and WMAcc_2bk correlation in non-zero pain
cor.test(SEM2_nonzero_painint$vmPFC, SEM2_nonzero_painint$WMAcc_2bk, method=c("pearson"))

#plot 
vmPFC_wmacc_plot<-ggplot(SEM2_nonzero_painint, aes(x=vmPFC, y=WMAcc_2bk,color="royalblue1")) +
  geom_point(size=2,color="royalblue1") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + xlab("vmPFC Activity") + ylab("2-Back Accuracy") + theme(legend.position = "none")

#combine into one plot
figure_ROIS <- ggarrange(aMCC_wmacc_plot, dMFC_wmacc_plot,vmPFC_wmacc_plot,
                    labels = c("b", "c", "d"),
                    ncol = 3, nrow = 1)
figure_ROIS

#correct p-values for multiple comparisons
#put unadjusted p-values into a vector
p<- c(0.8319,0.9868,0.0000897) 

#Bonferroni correct p-values
p.adjust(p, method = "bonferroni", n = length(p))

# aMCC=1.0000000
# dMFC=1.0000000
# vmPFC=0.0002691  ***

#------------------------------------------
#WMAcc_2bk and ListSort correlation in non-zero pain
cor.test(SEM2_nonzero_painint$ListSort_AgeAdj, SEM2_nonzero_painint$WMAcc_2bk, method=c("pearson"))

#plot 
ggplot(SEM2_nonzero_painint, aes(x=ListSort_AgeAdj, y=WMAcc_2bk,color="coral")) +
  geom_point(size=2,color="coral") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + xlab("List Sorting Score") + ylab("2-Back Accuracy") + theme(legend.position = "none")

#---------------------------------
#pain intensity and Age in Years in nonzero pain sample
cor.test(SEM2_nonzero_painint$Pain_Intensity, SEM2_nonzero_painint$Age_in_Yrs, method=c("pearson"))

#plot
ggplot(SEM2_nonzero_painint, aes(x=Age_in_Yrs, y=Pain_Intensity,color="royalblue1")) +
  geom_point(size=3,color="royalblue1") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + xlab("Age in Years") + ylab("Pain Intensity")

#WMacc_2bk and Age in Years full sample 
cor.test(SEM2_nonzero_painint$WMAcc_2bk, SEM2_nonzero_painint$Age_in_Yrs, method=c("pearson"))

#plot
ggplot(SEM2_nonzero_painint, aes(x=Age_in_Yrs, y=WMAcc_2bk,color="royalblue1")) +
  geom_point(size=3,color="royalblue1") + 
  geom_smooth(data=SEM2_nonzero_painint,method=lm, color="black") + theme_classic() + xlab("Age in Years") + ylab("2-Back Accuracy (% Correct)") 

```

#Pain Intensity Associated with 2-Back Task Performance 
```{r Pain Intensity Associated with 2-Back Task Performance, warnings=FALSE, message=FALSE, cache=FALSE}
#subset only model variables
SEM2_nonzero_painint_sub<-SEM2_nonzero_painint[c("Pain_Intensity","WMAcc_2bk","Sadness","Fear","Anger","Stress","aMCC","dMFC","vmPFC")]

#specify model with WMAcc_2bk outcome
painintnsmod2 <- '
neg =~ Anger + Stress + Sadness + Fear

#direct effect
WMAcc_2bk ~ c1*Pain_Intensity

#mediation a paths
aMCC ~ a1*Pain_Intensity
dMFC ~ a2*Pain_Intensity
vmPFC ~ a3*Pain_Intensity

neg ~ a4*Pain_Intensity

#mediation b paths
WMAcc_2bk ~ b1*aMCC
WMAcc_2bk ~ b2*dMFC
WMAcc_2bk ~ b3*vmPFC

WMAcc_2bk ~ b4*neg

#mediation d paths
aMCC ~ d1*neg
dMFC ~ d2*neg
vmPFC ~ d3*neg

#indirect effects
pain_aMCC_WM := a1*b1
pain_dMFC_WM := a2*b2
pain_vmPFC_WM := a3*b3
pain_neg_WM := a4*b4

 pain_neg_aMCC_WM := a4*d1*b1
 pain_neg_dMFC_WM := a4*d2*b2
 pain_neg_vmPFC_WM := a4*d3*b3

#mod indices
vmPFC ~~ dMFC
aMCC ~~ vmPFC
aMCC ~~ dMFC

#total effect
totpainintense := 
c1 + (a1*b1) + (a2*b2) + (a3*b3) + (a4*d1*d2*d3*b4)

'

#fit model
totfit2 <- lavaan::sem(painintnsmod2, data=SEM2_nonzero_painint_sub, estimator = "MLR")

#get fit indices and path estimates
summary(totfit2, fit.measures = TRUE, standardized=TRUE, modindices = TRUE)

#get parameter estimates
fit_pars<-parameterEstimates(totfit2)

#plot
semPaths(totfit2,what = "path",whatLabels="est",title=F,sizeMan=11,sizeLat=11,rotation=3,curvePivot = T,layout="tree",edge.label.cex=1,thresholds=TRUE,intercepts = FALSE,residuals=F,reorder=T)

#create table
#semTable(totfit2,type=c("csv","html"),file="//datastore01.psy.miami.edu/Groups/ELosin_Lab/Research/Research_Projects/SEM_PainCognition/Scripts/SEM_tableoutput_2back")

#run again without MLR and using bootstrapped indirect effects
medmodel_boostrapped_se <- lavaan::sem(painintnsmod2, data =SEM2_nonzero_painint_sub,se = "bootstrap", bootstrap = 1000)

summary(medmodel_boostrapped_se,standardized=TRUE,fit.measures=TRUE)

#---------------------------------
#specify model with only vmPFC 
painintnsmod2_vmpfc_only <- '
#direct effect
WMAcc_2bk ~ c1*Pain_Intensity

#mediation a paths
vmPFC ~ a1*Pain_Intensity

#mediation b paths
WMAcc_2bk ~ b1*vmPFC

#indirect effects
pain_vmPFC_WM := a1*b1

#total effect
totpainintense := c1 + (a1*b1) 
'

#fit model
totfit2_vmpfc_only <- lavaan::sem(painintnsmod2_vmpfc_only, data=SEM2_nonzero_painint_sub, estimator = "MLR")

#get fit indices and path estimates
summary(totfit2_vmpfc_only, fit.measures = TRUE, standardized=TRUE, modindices = TRUE)

#plot
semPaths(totfit2_vmpfc_only,what = "path",whatLabels="est",title=F,sizeMan=11,sizeLat=11,rotation=3,curvePivot = T,layout="tree",edge.label.cex=1,thresholds=TRUE,intercepts = FALSE,residuals=F,reorder=T)
```

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
